---
title: "Week 2 Assignment"
author: "Andrea Staten"
date: "8/31/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this assignment, we are going to use data from the National Assessment of Educational Progress (NAEP), or the "Nation's Report Card." The data here is for the 4th grade mathematics test. You can read more about it here: https://www.nationsreportcard.gov/

The NAEP is an important data source in education; read some recent coverage of scores (post-pandemic) here in USA Today: https://www.usatoday.com/story/news/education/2023/06/21/naep-scores-basic-reading-math-13-year-olds/70337312007/

## Step 1

Load necessary packages (you may need to install these packages first using `install.packages("___")`): rmarkdown, tidyverse, janitor, and skimr. Do so in the code chunk below. Be sure to comment (add a `#` as the first character) these lines after you have installed the packages; otherwise, R Markdown may run into issues when you try to knit the file.

```{r}
# Install packages below
#install.packages("rmarkdown")
#install.packages("tidyverse")
#install.packages("janitor")
#install.packages("skimr")
#install.packages("sf")

library(readr)
library(rmarkdown)
library(tidyverse)
library(janitor)
library(skimr)
```

## Step 2

Load in data. *Notes*: you must assign a name to the dataframe before loading it. Ensure that the data is saved _in the same folder_ as this file is saved.

```{r}
naep_data <- read_csv("naep-data.csv")
```

## Step 3

View your data. There are multiple ways to do this. Here are two options, with some code to get you started.

```{r}
glimpse(naep_data)

##commented out to avoid errors when knitting.
#view(naep_data)
```

## Step 4

Clean variable names using the janitor package.  You'll need to use the assignment operator to ensure your clean names save to a dataframe. You might consider using a name that includes the phrase "clean" to clarify to you and others what the difference is! 

Before doing so, take a peak at some of the handy functions that janitor provides, focusing on the `clean_names()` example: https://cran.r-project.org/web/packages/janitor/vignettes/janitor.html

```{r}
naep_clean <- naep_data %>% 
  janitor::clean_names()
```

## Step 5

Find the mean, standard deviation, quartiles, and more with the `skim()` function and the name of your data frame. Please first read a bit about skimr - how it works and what its benefits are - here: https://cloud.r-project.org/web/packages/skimr/vignettes/skimr.html

```{r}
naep_clean %>% 
  skim()
```

Congrats! You're on your way to exploring an authentic data set in R. Onto some questions:

## Step 6 (Reach)

Here's one optional step. Create a plot using either _base R_ or _ggplot2_. Use the code chunk below and add a bullet point or two interpreting your plot below the code you used to create it.

```{r}
library(sf)

states_sf <- st_read("US_State_Boundaries/US_State_Boundaries.shp")
states_sf <- states_sf %>% 
  rename(jurisdiction = NAME)

library(dplyr)
states_scores_sf <- left_join(states_sf, naep_clean, by = "jurisdiction")

library(ggplot2)
ggplot(states_scores_sf) +
  geom_sf(aes(fill = average_score_0_500)) +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(title = "Scores by State")

## This code is taking a United States shapefile I downloaded and joining the data to this shapefile by the state names! I then take the average score and plot that on a map using the shapefile!
```

## Questions

1.  What is the average score for the nation?

-   234.35294

2.  What is the average score for your home state?

-   236

3.  How does your home state and the nation compare?

-   My home state and the nation are comparable with a difference of less than 2.

4.  What state has the maximum and what is the value? What about the minimum?

-   Wyoming has the maxiumum at 243, while New Mexico has the minimum at 221.

## Rubric

|                     | Points Possible | Points Earned |
|---------------------|-----------------|---------------|
| File knits          | 5               |               |
| Code is appropriate | 5               |               |
| Question 1 answered | 5               |               |
| Question 2 answered | 5               |               |
| Question 3 answered | 5               |               |
| Question 4 answered | 5               |               |

## Knit and Submit

Complete the following steps to knit and publish your work:

1.  First, change the name of the `author:` in the [YAML header](https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html#yaml-metadata) at the very top of this document to your name. The YAML header controls the style and feel for knitted document but doesn't actually display in the final output.

2.  Next, click the knit button in the toolbar above to "knit" your R Markdown document to a [HTML](https://bookdown.org/yihui/rmarkdown/html-document.html) file that will be saved in your R Project folder. You should see a formatted webpage appear in your Viewer tab in the lower right pan or in a new browser window. Let's us know if you run into any issues with knitting.

3. Upload the HTML file that is now present in the same folder as this R Markdown file to Canvas to receive credit for this assignment!